---
title: "In Class Lab"
author: "Damon Ross"
date: "DATA 2401"
output: html_document
---
**Note: Your exam will be in a similar format to this (RMarkdown File with Questions and Code chunks for your answers), so make sure you have the process for knitting to html working.** 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## In Class Activity: 

1. Write a function to simulate rolling a 6-sided die and returns the result.

```{r}
die_roll1 <- function(){
  roll_1 <- sample(1:6, 1, replace = TRUE)
  
  return(roll_1)
}

die_roll1()
```



1. Write a function to simulate rolling 2 6-sided dice that returns the sum of the rolls. 

```{r}
die_roll2 <- function(){
  roll_1 <- sample(1:6, 1, replace = TRUE) #die 1 roll
 
  roll_2 <- sample(1:6, 1, replace = TRUE) #die 2 roll
  
  sum_roll <- paste("Die #1 shows:",roll_1, "and Die #2 shows:", roll_2, "and the sum is:", roll_1 + roll_2)
  
  return(sum_roll)
}

die_roll2()

```


1. Write a function that takes two people name's as strings, simulates rolling a 6-sided die for each of the names, then returns the name of the person with the higher roll. The function should reroll once on a tie, but if there is a second tie, it should return "draw". 

```{r}
highest_roll <- function(name1, name2) {
  
  name1_roll <- sample(1:6, 1, replace = TRUE) #die 1 roll
 
  name2_roll <- sample(1:6, 1, replace = TRUE) #die 2 roll
  
  if(name1_roll == name2_roll) {
      name1_roll <- sample(1:6, 1, replace = TRUE) #die 1 re-roll
      name2_roll <- sample(1:6, 1, replace = TRUE) #die 2 re-roll
      
  }
  
  if(name1_roll == name2_roll) {
      paste(name1_roll, name2_roll)
      return(print("No winner, it's a draw"))
      
  } else if(name1_roll > name2_roll) {
    return(paste(name1, "is the winner"))
    
  } else if(name1_roll < name2_roll) {
    return(paste(name2, "is the winner"))
    
  }  
  
} 

highest_roll("Jack", "Jill")

```


2. Write a second function (you may start by modifying the first one) that takes two names as strings, simulates both of them rolling die 15 times, and returns the name of the person who "won" more rolls. Don't reroll ties this time, but don't count them as "wins". 

```{r}
roll_15 <- function(name1, name2) {
  
  name1_roll <- sample(1:6, 15, replace = TRUE) #die 1 roll 15 times
 
  name2_roll <- sample(1:6, 15, replace = TRUE) #die 2 roll 15 times

  print(name1)
  print(name1_roll)
  
  print(name2)
  print(name2_roll)
  
  name1_win_total <- (sum(name1_roll > name2_roll))
  name2_win_total <- (sum(name2_roll > name1_roll))
  
  if(name1_win_total == name2_win_total) {
    return("Result is a draw")
    
  } else if(name1_win_total > name2_win_total) {
    return(paste(name1, "wins"))
    
  } else if(name2_win_total > name1_win_total) {
    return(paste(name2, "wins"))
  }
  
}

roll_15("Bob", "Sue")

```


3. Modify your function from (2) so that the person in the first argument has a die that is weighted to get higher numbers more often (you may decide the actual percentage change). Run the function ten or so times. Does the second person ever win?

```{r}
# player 1 gets weighted die for higher roll amoumts
roll_10 <- function(name1, name2) {
  
  name1_roll <- sample(1:6, 10, replace = TRUE, prob = c(2/20,2/20,2/20,3/20,5/20,6/20)) #die 1 roll 10 times
 
  name2_roll <- sample(1:6, 10, replace = TRUE) #die 2 roll 10 times

  print(name1)
  print(name1_roll)
  
  print(name2)
  print(name2_roll)
  
  name1_win_total <- (sum(name1_roll > name2_roll))
  name2_win_total <- (sum(name2_roll > name1_roll))
  
  if(name1_win_total == name2_win_total) {
    return("Result is a draw")
    
  } else if(name1_win_total > name2_win_total) {
    return(paste(name1, "wins"))
    
  } else if(name2_win_total > name1_win_total) {
    return(paste(name2, "wins"))
  }
  
}

roll_10("Bob", "Sue")
```

The 2nd person won the game about 20% of the time. 
